---
title: "NPS_Summary"
author: "ODONNELL"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, import needed libraries
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
```

Now, load the data into a dataframe
And, view the first few rows
```{r}
NPS_summary <- read_csv("NPS_Summary.csv")
NPS_summary$Description <- enc2utf8(NPS_summary$Description)
head(NPS_summary)
```

Now, check the dims and data types in the dataframe
```{r}
dim(NPS_summary)
str(NPS_summary)
```

Change one of the column names..
```{r}
colnames(NPS_summary)
colnames(NPS_summary)[colnames(NPS_summary)=="National Park"] <- "National_Park"
```

Lets add a columnn to show visitation increase %
```{r}
NPS_summary <- mutate(NPS_summary, Visitation_Increase = ((Visitors_2017-(rowMeans(NPS_summary[c(11:18)])))/(rowMeans(NPS_summary[c(11:18)])))*100)
```

OK, now lets see an overview of parks on the rise
caviat, parks with at least  1 Million visitors in 2017
```{r}
filter(NPS_summary, Visitors_2017 > 1000000) %>%
  group_by(National_Park, Visitation_Increase) %>%
  summarise(Visitors_Rank) %>%
  arrange(desc(Visitation_Increase))
```

And also, lets check how many national parks are in each state.
```{r}
NPS_summary %>%
  group_by(State) %>%
  summarise(NPs = n()) %>%
  arrange(desc(NPs))
```

And how about number of national park land (acres) by state!
```{r}
NPS_summary %>%
  group_by(State) %>%
  summarise(Land = sum(Acres)) %>%
  arrange(desc(Land))
```

To see which parks had highest tent campers per visitor by park
```{r}
NPS_summary %>%
  group_by(National_Park, State, Tent_Campers_2017) %>%
  summarise(TentCamp_Percentage = Tent_Campers_2017/Visitors_2017) %>%
  arrange(desc(TentCamp_Percentage))
```

To see which parks had highest campers:visitors by state
```{r}
NPS_summary %>%
  group_by(State, Tent_Campers_2017) %>%
  summarise(TentCamp_Percentage = sum(Tent_Campers_2017)/sum(Visitors_2017)) %>%
  arrange(desc(TentCamp_Percentage))
```

To see which state has the most National Parks Visitors in 2017
```{r}
NPS_summary %>%
  group_by(State) %>%
  summarise(total_visitors = sum(Visitors_2017)) %>%
  arrange(desc(total_visitors))
```

Which state has the highest National Park quality rating?
```{r}

NPS_summary %>%
  group_by(State) %>%
  summarise(NP_quality_score = mean(Quality_Rating)) %>%
  arrange(desc(NP_quality_score))
```

Let's shorten the Description column
to only the first sentance.
```{r}
NPS_summary$Desc <- str_extract(NPS_summary$Description, "((?<![[:space:]])[[:upper:]]).{20,}(?=\\.)")
```

Ok, lastly lets clean up the file to export as a CSV
```{r}
NPS_summary <- subset(NPS_summary, select = -c(Visitors_2009, Visitors_2010,
                                               Visitors_2011, Visitors_2012,
                                               Visitors_2013, Visitors_2014,
                                               Visitors_2015, Visitors_2016,
                                               Description))

write.csv(NPS_summary, "NPS_Summary_Tidy.csv")
```

File Summary:

1. Imported the data
2. Created ("Mutated") a column to display "Visitation Increase %"
3. Grouped and Summarised the data with Tidyr to analyze data by National Park and by State
4. Cleaned the "Description" column to only display the first sentance (of wikipedia description)
5. Subsetted the data and exported as a nice CSV
