---
title: "NPS_Details"
author: "ODONNELL"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, import needed libraries
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
```

Now, load the data into a dataframe
And, view the first few rows
```{r}
NPS_details <- read_csv("NPS_details_V2.csv")
head(NPS_details)
```

Now, check the dims and data types in the dataframe
```{r}
dim(NPS_details)
str(NPS_details)
```

Delete the unneeded columns
```{r}
NPS_details <- subset(NPS_details, select = -c(Keywords, Recreation,
                                               Flora, Fauna,
                                               Description))
```


Change one of the column names..
```{r}
colnames(NPS_details)[colnames(NPS_details)=="National Park"] <- "National_Park"
```

```{r}
NPS_details$All_Keywords <- tolower(NPS_details$All_Keywords)
```


Define keywords to go with recreational activities/attractions
```{r}
mountain_keywords <- c("mountain", "mountains")
ocean_keywords <- c("ocean", "coastline")
hiking_keywords <- c("hiking", "hike", "trails", "trailheads")
large_animal_keywords <- c("bears", "moose")
large_trees_keywords <- c("redwood", "redwoods", "giant sequoia", "douglas fur")
desert_keywords <- c("desert")
camping_keywords <- c("camping", "campers")
lake_keywords <- c("lake")
```

Create columns in the dataframe associated with
some attractions of national parks
```{r}
NPS_details$mountains <- 0
NPS_details$ocean <- 0
NPS_details$lake <- 0
NPS_details$desert <- 0
NPS_details$large_trees <- 0
NPS_details$hiking <- 0
NPS_details$camping <- 0
NPS_details$large_animals <- 0
```

Populate the columns just created above
```{r}
NPS_details$mountains <- str_detect(NPS_details$All_Keywords, mountain_keywords)
NPS_details$ocean <- str_detect(NPS_details$All_Keywords, ocean_keywords)
NPS_details$lake <- str_detect(NPS_details$All_Keywords, lake_keywords)
NPS_details$desert <- str_detect(NPS_details$All_Keywords, desert_keywords)
NPS_details$large_trees <- str_detect(NPS_details$All_Keywords, large_trees_keywords)
NPS_details$hiking <- str_detect(NPS_details$All_Keywords, hiking_keywords)
NPS_details$camping <- str_detect(NPS_details$All_Keywords, camping_keywords)
NPS_details$large_animals <- str_detect(NPS_details$All_Keywords, large_animal_keywords)
```

tidy the dataset
```{r}
#NPS_details[NPS_details==TRUE] <- colnames(NPS_details)[which(NPS_details==TRUE, arr.ind=TRUE)[,'col']]
#NPS_details[NPS_details==FALSE] <- colnames(NPS_details)[which(NPS_details==FALSE, arr.ind=FALSE)[,'col']]
#NPS_details <- gather(NPS_details, key = "Features", value = "n", 5:7)
```

Now, to analyze the dataset
Lets start by finding the frequency of best season
```{r}
NPS_details %>%
  group_by(Best_Season) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
```

Now that we know Summer has the most parks in the best season
lets filter for summer and check for parks with camping!
```{r}
filter(NPS_details, camping == "TRUE") %>%
  group_by(National_Park, State) %>%
  summarise(mountains) %>%
  arrange(desc(mountains))
```

And lastly, lets see which state is best for summer camping
at a National Park
```{r}
filter(NPS_details, camping == "TRUE") %>%
  group_by(State) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
```

Lastly, export as a CSV
```{r}
write.csv(NPS_details, "NPS_Details_Tidy.csv")
```

File Summary:

1. Imported the data
2. Deleted unneeded columns and added new columns
3. Used Stringr to go through wikipedia descriptions of each
National Park and pick out national park attractions
4. Used the new columns to group and summarise
National Parks and States to analyze them by Best Season
5. Learned that California is the best state to camp in
a National Park during the Summer
6. Exported the file as a CSV

3. Grouped and Summarised the data with Tidyr to analyze data by National Park and by State
4. Cleaned the "Description" column to only display the first sentance (of wikipedia description)
5. Subsetted the data and exported as a nice CSV