---
title: "NPS POI's"
author: "ODONNELL"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, import needed libraries
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
```

Now, load the data into a dataframe
And, view the first few rows
```{r}
NPS_POIs <- read_csv("NPS_POIs_V2.csv")
head(NPS_POIs)
```

Now, check the dims and data types in the dataframe
```{r}
dim(NPS_POIs)
str(NPS_POIs)
```

OK, there's quite a few columns we do not need
Therefore, let's remove them
```{r}
NPS_POIs <- subset(NPS_POIs, select = -c(POISTATUS, PUBLICDISPLAY,
                                         DATAACCESS, ACCESSNOTES,
                                         ORIGINATOR, UNITTYPE,
                                         GROUPCODE, GROUPNAME,
                                         MAPMETHOD, MAPSOURCE,
                                         SOURCEDATE, XYACCURACY,
                                         GEOMETRYID, FEATUREID,
                                         FACLOCID, FACASSETID,
                                         IMLOCID, OBSERVABLE,
                                         ALTLANGNAME, ALTLANG,
                                         StagingTable, POIALTNAME))
```

Much easier to read now
To continue improvement, lets rename some columns
```{r}
colnames(NPS_POIs)[colnames(NPS_POIs)=="X"] <- "Longitude"
colnames(NPS_POIs)[colnames(NPS_POIs)=="Y"] <- "Latitude"
colnames(NPS_POIs)[colnames(NPS_POIs)=="POITYPE"] <- "POI_Type"
colnames(NPS_POIs)[colnames(NPS_POIs)=="UNITNAME"] <- "National_Park"
colnames(NPS_POIs)[colnames(NPS_POIs)=="SEASDESC"] <- "Seasonal_Desc"
colnames(NPS_POIs)
```

Alright, so which National Parks do we have in this dataset?
```{r}
unique(NPS_POIs$National_Park)
```

Now, what type of POI's are in this dataset?
```{r}
unique(NPS_POIs$POI_Type)
```

Cool! lets find the most popular POI's
```{r}
NPS_POIs %>%
  group_by(POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N))
```

Neat! Alright lets summarize POI's by National Park!
```{r}
NPS_POIs %>%
  group_by(National_Park, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(National_Park, desc(N))
```

OK to keep digging, lets only look at POI_types related to hiking!
```{r}
filter(NPS_POIs, POI_Type %in% c("Trailhead",
                                 "")) %>%
  group_by(National_Park, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N))
```

To get a different picture, lets combine anything that
relates to hiking
```{r}
filter(NPS_POIs, POI_Type %in% c("Trailhead",
                                 "Trail Access",
                                 "Trail Marker")) %>%
  group_by(National_Park, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) %>%
  spread(POI_Type, N) %>%
  mutate(Hiking_POIs = sum(Trailhead, `Trail Marker`, `Trail Access`))
```

Alright, let's do the same for camping!
This is getting interesting.
```{r}
filter(NPS_POIs, POI_Type %in% c("Campsite")) %>%
  group_by(National_Park, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N))
```

And lastly for visitor centers,
to get a sense of size
```{r}
filter(NPS_POIs, POI_Type %in% c("Visitor Center")) %>%
  group_by(National_Park, POI_Type) %>%
  summarise(N = sort(table(POI_Type))) %>%
  arrange(desc(N))
```

Maybe we can make some interesting conclusions by region.
```{r}
unique(NPS_POIs$REGIONCODE)
```

First lets just list all POI_Types
```{r}
NPS_POIs %>%
  group_by(REGIONCODE, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N))
```

Now, lets just look at camping and hiking POIs
```{r}
filter(NPS_POIs, POI_Type %in% c("Campsite", "Trailhead")) %>%
  group_by(REGIONCODE, POI_Type) %>%
  summarise(N = n()) %>%
  arrange(desc(N))
```

After analysis, I only want the following dataframe
```{r}
NPS_POIs <- NPS_POIs %>%
  group_by(National_Park, POI_Type) %>%
  summarise(Count = n()) %>%
  arrange(National_Park)
```


Lastly, export the file as a CSV
```{r}
write.csv(NPS_POIs, "NPS_POIs_Tidy.csv")
```

File summary:

1. Imported the data from a CSV
2. Subsetted the data to remove unneeded columns
3. Renamed columns to make analysis easier
4. Grouped and Summarised (with tidyr) the data to analyze the Points of Inerest at many levels
5. Made conclusions about POI's and National Parks
6. Subsetting the data for the most important infomation I found and exported as a CSV!
